Follow the Flow of Rain and Paint: \break An Output Sensitive Contour Tree Algorithm

### 2014-10-03 paint-contour-tree.tex
Section 5 updated

We introduce the concept of path decompositions. A \emph{leaf path} is simply a contiguous path in $T$
containing a leaf.
% Then in the following two subsections 
%we relate path decomposition costs (defined below) to the running time of our algorithm using $\Lem{algCost}$.

@@

\begin{proof} For convenience, denote $H_i = H_{q_i}$ and $H_0 = H_{\ell(p)}$.
Consider $v \in \bigcup_{i} H_i$ that lies below
$b(p)$ in $T$. Note that $v \in H_p$ and has only one color in $\bigcup_{i} L(T_(b(q_i))) \cup \ell(p)$.
Since the latter sets are all disjoint, any $u \in H_i \cap H_j$ must lie in $p$.
Any $u \in p$ can be in at most two of the $H_i$'s. Hence, $\sum_i |H_i| \leq |H_p| + 2|p|$.
\ignore{
 First observe that for all $i$, $r_{i}$ lies above $r_p$ on some root to leaf path. 
 Let $H_i'$ be the subset of $H_i$ that lies below $r_p$.  Observe that $H_i'\subseteq H_p$, as $L(T_{r_i})\subseteq L(T_{r_p})$.

@@

There are some light paths that we can also ``pay" for.

\begin{definition} \label{def:support} A light path $p$ is \emph{supported} if
at least $|H_p|/100$ vertices $v$ in $H_p$ lie in heavy ancestors.
%
%
% there exists
%a contiguous (from $p$) chain of ancestors $q_1, q_2, \ldots, q_k$ (in $P(T)$) such that $|H_{q_k}| \leq 99|H_p|/100$
%and all $q_i$'s are heavy.
\end{definition}

Let $\cL$ be the set of light paths, $\cL'$ be the set of supported light paths, and $\cH$ be the set of heavy paths
given by $\pmax(T)$.
We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. Hence, $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
to the shrub forest of $\cL \setminus \cL'$, where every unsupported light path is an internal node, and all
roots are supported light paths.

@@

We split the summation into heavy, light, and unsupported light paths.
\begin{eqnarray*} 
\sum_{p \in \cL} \sum_{v \in p} \log |H_v| & = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cH} \sum_{v \in p} \log |H_v|
%& = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |p|\log |H_p|) = \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |H_p|\log |H_p|)
\end{eqnarray*}
The last term can be bounded by $O(\sum_{p \in \pmax(T)} |p|\log |p|)$, by \Lem{pathBounds}.
The second term can be bounded by $O(\sum_{p \in \cL'} |H_p| \log |H_p|)$, by \Lem{pathBounds} again.
The following claim shows that in turn is at most the last term.

@@

\begin{proof} 
Pick $p \in \cL'$. 
%Let the first $|H_p|/200$ vertices the lie in heavy paths be denoted as $H'_p$. 
%Observe that for any $v \in H'_p$, 
%We have $|H_p| \log |H_p| = O(\sum_{v \in H'_p} \log |H_p|)$
%$= O(\sum_{v \in H'_p} \log |H_v|)$. Since all the vertices lie in heavy paths,
%we can write this as $O(\sum_{q \in \cH} \sum_{v \in H'_p \cap q} \log |H_v|)$.
%Summing over all $p$, the expression is $\sum_{q \in \cH} \sum_{p \in \cL'} \sum_{v \in H'_p \cap q} \log |H_v|$.
%
%
As we traverse the heavy ancestors, there
are at least $|H_p|/100$ vertices of $H_p$ that lie in these paths. These are encountered in
a fixed order. Let $H'_p$ be the first $|H_p|/200$ of these vertices. When $v \in H'_p$
is encountered, there are $|H_p|/200$ vertices of $H_p$ not yet encountered. Hence,
$|H_v| \geq |H_p|/200$. Hence, $|H_p|\log |H_p| = O(\sum_{v \in H'_p} \log |H_v|)$.
Since all the vertices lie in heavy paths,
we can write this as $O(\sum_{q \in \cH} \sum_{v \in H'_p \cap q} \log |H_v|)$.
Summing over all $p$, the expression is $\sum_{q \in \cH} \sum_{p \in \cL'} \sum_{v \in H'_p \cap q} \log |H_v|$.
%and $\sum_{p \in \cL'} |H_p|\log |H_p| = O(\sum_{p \in \cL'} \sum_{v \in H'_p} \log |H_v|)$.

Pick another $\widetilde{p} \in \cL'$. If $p$ and $\widetilde{p}$ are ancestor-descendant,
then their corresponding supporting ancestor chains are disjoint. Any $v \in H'_p$
can only occur in one other $H'_{\hat{p}}$ (for $\hat{p} \in \cL'$), because $v$ has exactly two colors.
In other words, any $\log |H_v|$ appears at most twice in the above triple summation.
Hence, we can bound it by $O(\sum_{q \in \cH} \sum_{v \in q} \log |H_v|)$.
%
%Pick $q \in Q$ and some $v \in H_q$. As we traverse up the tree passing
%through $A(q)$, we must encounter some $p \in A(q)$ such that $v \in p$. Furthermore,
%there is an order in which these vertices (of $H_q$) occur. Let $H'$ be the first half
%of vertices that occur. For all $v \in H'$, $|H_v| \geq |H_q|/2$. Hence,
%$|H_q| \log |H_q| \leq 4 \sum_{v \in H'} \log |H_v| \leq 4 \sum_{v \in H_q} |H_v|$.
%So $\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{q \in Q} \sum_{v \in H_q} |H_v|$.
%
%Since nodes in $Q$ are incomparable, all the $H_q$'s (for $q \in Q$) are disjoint.
%Each such $v$ must occur in some path $p \in A(Q)$. Thus,
%$\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{p \in A(Q)} \sum_{v \in p} |H_v|$.
%We apply the bound of \Lem{pathBounds} to complete the proof.
\end{proof}

@@

\begin{proof} Let $\cU$ denote a connected component of the shrub forest.
We have $\sum_{p \in \cL \setminus \cL'} \leq \sum_{\cU} \sum_{p \in \cU} \sum_{v \in p} \log |H_v|$.
By \Lem{cu-root}, $|H_v| = O(|H_r|)$, where $r$ is the root of $\cU$.
Furthermore, $\sum_{p \in \cU} |p| = O(|H_r|)$. 
We have $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O((\log |H_r|) \sum_{p \in \cU} |p|) = O(|H_r|\log |H_r|)$.
We sum this over all $\cU$ in the shrub forest, and note that roots in the shrub forest are supported light paths.
\end{proof}

@@

\Lem{cu-root} asserts the root $r$ in $\cU$ pretty much encompasses all sizes and heaps in $\cU$.
We will work with the \emph{reduced} heap $\redH_p$. This is the subset of $H_p$
of vertices that only appear in light paths. By definition, for any unsupported light path
(hence, any non-root $p \in \cU$), $|H'_p| \geq 99|H_p|/100$. For notational convenience, we simply
set $H'_r = H_r$.
%Consider $q \in \cL \setminus \cL'$ that is a child of $p$ in the shrub $\cU$.
We begin with a key property, which is where the construction of $\pmax(T)$ enters the picture.

@@

\begin{lemma}
\label{lem:geometric}
 Let $p$ be the parent of $q$ in the shrub $\cU$. Then $|\redH_p|\geq \frac{3}{2}|\redH_q|$. 
\end{lemma}
\begin{proof} By construction, there is a heavy path (denoted $h(q)$) that is a child of $p$ in $\pmax(T)$,
and an ancestor of $q$. The chain of ancestors from $q$ to $h(q)$ consists only of heavy paths.
Since $q$ is unsupported, these paths contain at most $|H_p|/100$ vertices of $H_p$. Thus, $|H_{h(q)}| \geq 99|H_p|/100$.

Consider the base
of $h(q)$, which is a node $w$ in $T$. Let $v$ denote the sibling of $w$ in $T$.
Their parent is called $u$. Note that both $u$ and $v$ are nodes in the path $p$.
Now, the decomposition $\pmax(T)$ put $u$ and $v$ in the same path $p$. This
implies $|H_v| \geq |H_w|$. Since $|H_u| \geq |H_v| + |H_w| - 1$,
$|H_u| \geq 2|H_w| - 1$. Let $b$ be the base of $p$.
We have $|H_p| = |H_b| \geq |H_u| - |p| \geq 2|H_w| - |p| - 2$. Since $p$ is a light path, $|p| < \sqrt{|H_p|}/100$.
Applying this bound, we get $|H_p| \geq (2-\delta)|H_w|$ (for small constant $\delta > 0$).
Since $w$ is the base of $h(q)$, $H_w = H_{h(q)}$. We apply the bound $|H_{h(q)}| > 99|H_q|/100$
to get $|H_p| \geq 197|H_q|/100$. The proof is completed by noting that $|\redH_p| \geq 99|H_p|/100$ and $|H_q| \geq |\redH_q|$.
%
%
% Let $w$ be the vertex in the path $p$ to which the root of the path $q$ is adjacent to in $T$, 
% i.e. $w$ is the parent of $r_q$ in $T$.  Let $c$ be the other child of $w$, which lies in $p$.
% As $\mathcal{P}(T)$ is a maximum path decomposition of $T$ and $c$ rather than $r_q$ is in 
% the same path as $w$, we have $|H_q|\leq |H_c|$.  Therefore, $|H_w|\geq |H_q|+|H_c|-1\geq 2|H_q|-1$. 
% Therefore, by \Obs{decrease}, $|H_p|+ d(r_p, w) \geq |H_w|\geq 2|H_q|-1$.  
% Since $p$ is a light path, $d(r_p, w)\leq \sqrt{|H_p|}/100$ and so we have $|H_p|+\sqrt{|H_p|}/100\geq 2|H_q|-1$.
% Assuming $|H_p|>11$ (and we can safely assume $|H_p|$ is bigger than any constant), this yields $\frac{3}{5} |H_p| \geq |H_q|$
\end{proof}

This immediately proves part (i) of \Lem{cu-root}. Part (ii) requires much more work.
It is helpful to spell out how the various heaps evolve. Consider $p \in \cU$,
which has children $q_1, q_2, \ldots, q_k$ in $\pmax(T)$.
Furthermore, as a leaf path in $p$, some leaf
$\ell$ of $T$ in one of the ends of $p$. The reduced heap $\redH_p$ is the
union of $\redH_\ell$ and $\redH_{q_1}, \redH_{q_2}, \ldots, \redH_{q_k}$, with the vertices in $p$
removed. Note that some of the $q_i$'s are the corresponding $h(q')$ vertices,
for a child $q'$ in $\cU$.

@@

\begin{claim} \label{clm:loss} $|\redH_p| \geq W_p - 2\sum_{q \in \cU_p} |q|$
\end{claim}

\begin{proof} We write $W_p - |\redH_p| = \sum_{q \in \cU_p} R^+_q - R_q$ $= -\sum_{q \in \cU_p: R_q < 0} R_q$.
Consider $q \in \cU_p$ and denote the children in $\cU_p$ by $q'_1, q'_2, \ldots$.
Note that $R_q$ is negative exactly when $|\redH_q| < \sum_i |\redH_{q'_i}|$.
Traverse $\pmax(T)$ from $q'_i$ to $q$.
Other than $q$, all other nodes encountered are heavy paths and hence
do not affect reduced heaps.
The vertices of $\redH_{q'_i}$ that are deleted
are exactly those present in the path $q$ (which are paths in $T$) traversed. 
Any vertex in $q$ can be deleted from at most two reduced heaps $\redH_{q'_i}$.
When $R_q$ is negative, it is at most $2|q|$.
We sum over all $q$ to complete the proof.
\end{proof}

@@

\begin{claim} \label{clm:charge} Fix $p \in \cU$. Suppose for any $q,q' \in \cU_p$
where $q$ is a parent of $q'$ in $\cU_p$, $W_q \geq (4/3)W_{q'}$. Then $\sum_{q \in \cU_p} |q| \leq W_p/20$.
\end{claim}

\begin{proof} Since $q$ is an unsupported light path, $|q| < \sqrt{|H_q|}/100 \leq \sqrt{|\redH_q|}/99 \leq \sqrt{|W_q|}/99$.
We prove that $\sum_{q \in \cU_p} \sqrt{|W_q|}/99 \leq W_p/20$
by a charge redistribution scheme. Assume that each $q \in \cU_p$
starts with $\sqrt{W_p}/99$ units of charge. We redistribute this charge over all
nodes in $\cU_p$, and then calculate the total charge. For $q \in \cU_p$, spread
its charge to all nodes in $\cU_q$ proportional to $R^+$ values. In other words,
give $(\sqrt{W_q}/99)\cdot(R^+_{q'}/W_q)$ units of charge to each $q' \in \cU_q$.

After the redistribution, let us compute the charge deposited at $q$. Every ancestor in $\cU_p$ 
$q = a_0, a_1, a_2, \ldots, a_k$ contributes to the charge at $q$. The charge is expressed in the following
equation. We use the assumption that $W_{a_i} \geq (4/3) W_{a_{i-1}}$ and hence $W_{a_i} \geq (4/3)^i$.
$$ ({R^+_q}/99)\sum_{a_i} 1/\sqrt{W_{a_i}} \leq ({R^+_q}/99)\sum_{a_i} (3/4)^{i/2} \leq R^+_q/20 $$
The total charge is $\sum_{q \in \cU_p} R^+_p/20 = W_p/20$. 
\end{proof}

@@

\begin{proof} We use $r$ for $r(\cU)$. Our aim is to simply apply \Clm{charge} from $r(\cU)$. 
Suppose $W_q \geq (4/3)W_{q'}$ (for any choice in $\cU$ of $q$ parent of $q'$). By \Clm{charge}, $\sum_{p \in \cU} |p| \leq W_{r}/20$.
By \Clm{loss}, $|\redH_r| \geq W_r - 2\sum_{p \in \cU} |p|$. By simple algebra, we deduce, $\sum_{p \in \cU} |p| \leq |\redH_r|/10 = |H_r|/10$.

We now prove that for any $q$ parent of $q'$, $W_q \geq (4/3)W_{q'}$.
Suppose not. Let $p, p'$ be the counterexample furthest from the root,
where $p$ is the parent of $p'$. 
Note that for $q$ and child $q'$ in $\cU_{p'}$, $W_q \geq (4/3)W_{q'}$.
We will apply \Clm{charge} for $\cU_{p'}$ to deduce that $\sum_{q \in \cU_{p'}} |q| \leq W_{p'}/20$.
By \Clm{loss}, $|\redH_{p'}| \geq 19W_{p'}/20$. By \Lem{geometric}, $|\redH_p| \geq (3/2)|\redH_{p'}|$.
Noting that $W_p \geq |\redH_p|$, we deduce that $W_p \geq (4/3)W_{p'}$.
Hence, $p, p'$ is not a counterexample, and more generally, this is no counterexample.
That completes the whole proof.
\end{proof}

>>>>

We introduce the concept of path decompositions. A \emph{leaf path} is simply a contiguous path in $T$
containing a leaf, which is monotonic in the height values of its vertices.
% Then in the following two subsections 
%we relate path decomposition costs (defined below) to the running time of our algorithm using $\Lem{algCost}$.

@@

\begin{proof} For convenience, denote $H_i = H_{q_i}$ and $H_0 = H_{\ell(p)}$.
Consider $v \in \bigcup_{i} H_i$ that lies below $b(p)$ in $T$. 
Note that such a vertex has only one of its two colors in $L(b(p))$.  
Since the colors tracked by $H_i$ and $H_j$ for $i\neq j$ are disjoint, 
such a vertex can appear in only one of the $H_i$'s.
On the other hand, a vertex $u\in p$ can appear in more than one $H_i$, 
but since any vertex has exactly two colors it can appear in at most two 
such heaps.  Hence, $\sum_i |H_i| \leq |H_p| + 2|p|$.
\ignore{
 First observe that for all $i$, $r_{i}$ lies above $r_p$ on some root to leaf path. 
 Let $H_i'$ be the subset of $H_i$ that lies below $r_p$.  Observe that $H_i'\subseteq H_p$, as $L(T_{r_i})\subseteq L(T_{r_p})$.

@@

There are some light paths that we can also ``pay" for.  Consider any light path $p$ in the shrub.  We will refer to the 
\emph{heavy support chain} of $p$ as the heavy ancestors of $p$ in the shrub which have a path to $p$ which does not use any 
light path (i.e. it is a chain of paths adjacent to $p$).

\begin{definition} \label{def:support} A light path $p$ is \emph{supported} if
at least $|H_p|/100$ vertices $v$ in $H_p$ lie in paths in the heavy support chain of $p$.
%
%
% there exists
%a contiguous (from $p$) chain of ancestors $q_1, q_2, \ldots, q_k$ (in $P(T)$) such that $|H_{q_k}| \leq 99|H_p|/100$
%and all $q_i$'s are heavy.
\end{definition}

Let $\cL$ be the set of light paths, $\cL'$ be the set of supported light paths, and $\cH$ be the set of heavy paths
given by $\pmax(T)$.
We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. Hence, $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
to the shrub forest of $\cL \setminus \cL'$, where all the roots are supported light paths, 
and the remaining nodes are the unsupported light paths.

@@

We split the summation into heavy, light, and unsupported light paths.
\begin{eqnarray*} 
\sum_{p \in \cL} \sum_{v \in p} \log |H_v| & = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cH} \sum_{v \in p} \log |H_v|
%& = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |p|\log |H_p|) = \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |H_p|\log |H_p|)
\end{eqnarray*}
The last term can be bounded by $O(\sum_{p \in \pmax(T)} |p|\log |p|)$, by \Lem{pathBounds}.
The second term can be bounded by $O(\sum_{p \in \cL'} |H_p| \log |H_p|)$, by \Lem{pathBounds} again.
The following claim shows that this in turn is at most the last term.

@@

\begin{proof} 
Pick $p \in \cL'$. 
%Let the first $|H_p|/200$ vertices the lie in heavy paths be denoted as $H'_p$. 
%Observe that for any $v \in H'_p$, 
%We have $|H_p| \log |H_p| = O(\sum_{v \in H'_p} \log |H_p|)$
%$= O(\sum_{v \in H'_p} \log |H_v|)$. Since all the vertices lie in heavy paths,
%we can write this as $O(\sum_{q \in \cH} \sum_{v \in H'_p \cap q} \log |H_v|)$.
%Summing over all $p$, the expression is $\sum_{q \in \cH} \sum_{p \in \cL'} \sum_{v \in H'_p \cap q} \log |H_v|$.
%
%
As we traverse the heavy support chain of $p$, there
are at least $|H_p|/100$ vertices of $H_p$ that lie in these paths. These are encountered in
a fixed order. Let $H'_p$ be the first $|H_p|/200$ of these vertices. When $v \in H'_p$
is encountered, there are $|H_p|/200$ vertices of $H_p$ not yet encountered. Hence,
$|H_v| \geq |H_p|/200$. Hence, $|H_p|\log |H_p| = O(\sum_{v \in H'_p} \log |H_v|)$.
Since all the vertices lie in heavy paths,
we can write this as $O(\sum_{q \in \cH} \sum_{v \in H'_p \cap q} \log |H_v|)$.
Summing over all $p$, the expression is $\sum_{q \in \cH} \sum_{p \in \cL'} \sum_{v \in H'_p \cap q} \log |H_v|$.
%and $\sum_{p \in \cL'} |H_p|\log |H_p| = O(\sum_{p \in \cL'} \sum_{v \in H'_p} \log |H_v|)$.

Consider any $v \in H'_p$.  Let $S$ be the set of paths $\widetilde{p}\in \cL'$ such that 
$v\in H'_{\widetilde{p}}$.  We now show $|S|\leq 2$ (i.e. it contains at most one path other than $p$).
First observe that any two paths in $S$ must be unrelated (i.e. $S$ is an anti-chain), 
since paths which have an ancestor-descendant relationship have disjoint heavy support chains.  
However, any vertex $v$ receives exactly one color from each of its two subtrees (in $T$), and therefore $|S|\leq 2$ since any two 
paths which share descendant leaves in $T$ (i.e. their heaps are tracking the same color) must have an ancestor-descendant relationship.

In other words, any $\log |H_v|$ appears at most twice in the above triple summation.
Hence, we can bound it by $O(\sum_{q \in \cH} \sum_{v \in q} \log |H_v|)$.
%
%Pick $q \in Q$ and some $v \in H_q$. As we traverse up the tree passing
%through $A(q)$, we must encounter some $p \in A(q)$ such that $v \in p$. Furthermore,
%there is an order in which these vertices (of $H_q$) occur. Let $H'$ be the first half
%of vertices that occur. For all $v \in H'$, $|H_v| \geq |H_q|/2$. Hence,
%$|H_q| \log |H_q| \leq 4 \sum_{v \in H'} \log |H_v| \leq 4 \sum_{v \in H_q} |H_v|$.
%So $\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{q \in Q} \sum_{v \in H_q} |H_v|$.
%
%Since nodes in $Q$ are incomparable, all the $H_q$'s (for $q \in Q$) are disjoint.
%Each such $v$ must occur in some path $p \in A(Q)$. Thus,
%$\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{p \in A(Q)} \sum_{v \in p} |H_v|$.
%We apply the bound of \Lem{pathBounds} to complete the proof.
\end{proof}

@@

\begin{proof} Let $\cU$ denote a connected component of the shrub forest.
We have $\sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| \leq \sum_{\cU} \sum_{p \in \cU} \sum_{v \in p} \log |H_v|$.
By \Lem{cu-root}, $|H_v| = O(|H_r|)$, where $r$ is the root of $\cU$.
Furthermore, $\sum_{p \in \cU} |p| = O(|H_r|)$. 
We have $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O((\log |H_r|) \sum_{p \in \cU} |p|) = O(|H_r|\log |H_r|)$.
We sum this over all $\cU$ in the shrub forest, and note that roots in the shrub forest are supported light paths.
\end{proof}

@@

\Lem{cu-root} asserts the root $r$ in $\cU$ pretty much encompasses all sizes and heaps in $\cU$.
We will work with the \emph{reduced} heap $\redH_p$. This is the subset of vertices of $H_p$ that 
do not appear on the heavy support chain of $p$.
%only appear in light paths. 
By definition, for any unsupported light path
(hence, any non-root $p \in \cU$), $|\redH_p| \geq 99|H_p|/100$. 
% For notational convenience, we simply set $\redH_r = H_r$.
%Consider $q \in \cL \setminus \cL'$ that is a child of $p$ in the shrub $\cU$.
We begin with a key property, which is where the construction of $\pmax(T)$ enters the picture.

@@

\begin{lemma}
\label{lem:geometric}
 Let $q$ be the child of some path $p$ in $\cU$, then $|H_p|\geq \frac{3}{2}|H_q|$. 
 Moreover, if $p\neq r(\cU)$, then $|\redH_p|\geq \frac{3}{2}|\redH_q|$. 
\end{lemma}
\begin{proof} Let $h(q)$ denote the heavy path that is a child of $p$ in $\pmax(T)$,
and an ancestor of $q$. If no such heavy path exists, then by construction $p$ is the parent 
of $q$ in $\pmax(T)$, and the following argument will go through by setting $h(q)=q$.

The chain of ancestors from $q$ to $h(q)$ consists only of heavy paths.
Since $q$ is unsupported, these paths contain at most $|H_q|/100$ vertices of $H_q$. 
Thus, $|H_{h(q)}| \geq 99|H_q|/100$.

Consider the base
of $h(q)$, which is a node $w$ in $T$. Let $v$ denote the sibling of $w$ in $T$.
Their parent is called $u$. Note that both $u$ and $v$ are nodes in the path $p$.
Now, the decomposition $\pmax(T)$ put $u$ and $v$ in the same path $p$. This
implies $|H_v| \geq |H_w|$. Since $|H_u| \geq |H_v| + |H_w| - 2$,
$|H_u| \geq 2|H_w| - 2$. Let $b$ be the base of $p$.
We have $|H_p| = |H_b| \geq |H_u| - |p| \geq 2|H_w| - |p| - 2$. Since $p$ is a light path, $|p| < \sqrt{|H_p|}/100$.
Applying this bound, we get $|H_p| \geq (2-\delta)|H_w|$ (for a small constant $\delta > 0$).
Since $w$ is the base of $h(q)$, $H_w = H_{h(q)}$. We apply the bound $|H_{h(q)}| \geq 99|H_q|/100$
to get $|H_p| \geq 197|H_q|/100$, implying the first part of the lemma.  For the second part, observe that 
if $p\neq r(\cU)$, then $p$ is unsupported and so $|\redH_p| \geq 99|H_p|/100$, and therefore the second part follows since $|H_q| \geq |\redH_q|$.
%
%
% Let $w$ be the vertex in the path $p$ to which the root of the path $q$ is adjacent to in $T$, 
% i.e. $w$ is the parent of $r_q$ in $T$.  Let $c$ be the other child of $w$, which lies in $p$.
% As $\mathcal{P}(T)$ is a maximum path decomposition of $T$ and $c$ rather than $r_q$ is in 
% the same path as $w$, we have $|H_q|\leq |H_c|$.  Therefore, $|H_w|\geq |H_q|+|H_c|-1\geq 2|H_q|-1$. 
% Therefore, by \Obs{decrease}, $|H_p|+ d(r_p, w) \geq |H_w|\geq 2|H_q|-1$.  
% Since $p$ is a light path, $d(r_p, w)\leq \sqrt{|H_p|}/100$ and so we have $|H_p|+\sqrt{|H_p|}/100\geq 2|H_q|-1$.
% Assuming $|H_p|>11$ (and we can safely assume $|H_p|$ is bigger than any constant), this yields $\frac{3}{5} |H_p| \geq |H_q|$
\end{proof}

This immediately proves part (i) of \Lem{cu-root}. Part (ii) requires much more work.
% It is helpful to spell out how the various heaps evolve. Consider $p \in \cU$,
% which has children $q_1, q_2, \ldots, q_k$ in $\pmax(T)$, and a leaf vertex $\ell$ in $T$. 
% The reduced heap $\redH_p$ is the
% union of $\redH_\ell$ and $\redH_{q_1}, \redH_{q_2}, \ldots, \redH_{q_k}$, with the vertices in $p$
% removed. Note that some of the $q_i$'s are the corresponding $h(q')$ vertices,
% for a child $q'$ in $\cU$.

@@

\begin{claim} \label{clm:loss} For any path $p\in \cU$, $|\redH_p| \geq W_p - 2\sum_{q \in \cU_p} |q|$.
\end{claim}

\begin{proof} We write $W_p - |\redH_p| = \sum_{q \in \cU_p} R^+_q - R_q$ $= -\sum_{q \in \cU_p: R_q < 0} R_q$.
Consider $q \in \cU_p$ and denote the children in $\cU_p$ by $q'_1, q'_2, \ldots$.
Note that $R_q$ is negative exactly when $|\redH_q| < \sum_i |\redH_{q'_i}|$.
Traverse $\pmax(T)$ from $q'_i$ to $q$.
Other than $q$, all other nodes encountered are in the heavy support chain of $q'_i$ and hence
do not affect its reduced heap.
The vertices of $\redH_{q'_i}$ that are deleted
are exactly those present in the path $q$. 
Any vertex in $q$ can be deleted from at most two of the reduced heaps (of the children of $q$ in $\cU_p$), since 
theses reduced heaps do not have an ancestor-descendant relationship.
Therefore when $R_q$ is negative, it is at most by $2|q|$.
We sum over all $q$ to complete the proof.
\end{proof}

@@

\begin{claim} \label{clm:charge} Fix any path $p \in \cU\setminus\{r(\cU)\}$. Suppose for any $q,q' \in \cU_p$
where $q$ is a parent of $q'$ in $\cU_p$, $W_q \geq (4/3)W_{q'}$. Then $\sum_{q \in \cU_p} |q| \leq W_p/20$.
\end{claim}

\begin{proof} Since $q$ is an unsupported light path, $|q| < \sqrt{|H_q|}/100 \leq \sqrt{|\redH_q|}/99 \leq \sqrt{|W_q|}/99$.
We prove that $\sum_{q \in \cU_p} \sqrt{|W_q|}/99 \leq W_p/20$
by a charge redistribution scheme. Assume that each $q \in \cU_p^-$
starts with $\sqrt{W_q}/99$ units of charge. We redistribute this charge over all
nodes in $\cU_q$, and then calculate the total charge. For $q \in \cU_p$, spread
its charge to all nodes in $\cU_q$ proportional to $R^+$ values. In other words,
give $(\sqrt{W_q}/99)\cdot(R^+_{q'}/W_q)$ units of charge to each $q' \in \cU_q$.

After the redistribution, let us compute the charge deposited at $q$. Every ancestor in $\cU_p^-$ 
$q = a_0, a_1, a_2, \ldots, a_k$ contributes to the charge at $q$. The charge is expressed in the following
equation. We use the assumption that $W_{a_i} \geq (4/3) W_{a_{i-1}}$ and hence $W_{a_i} \geq (4/3)^i W_{a_0} \geq (4/3)^i$, 
as $a_0$ is an unsupported light path and hence $W_{a_0}\geq 1$. 
$$ ({R^+_q}/99)\sum_{a_i} 1/\sqrt{W_{a_i}} \leq ({R^+_q}/99)\sum_{a_i} (3/4)^{i/2} \leq R^+_q/20 $$
The total charge is $\sum_{q \in \cU_p} R^+_p/20 = W_p/20$. 
\end{proof}

\begin{corollary} \label{cor:chargeCor}
 Let $r$ be the root of $\cU$, and suppose that for any paths $q,q' \in \cU\setminus \{r\}$, where $q$ is a parent of $q'$ in $\cU$, $W_q \geq (4/3)W_{q'}$.
 Then $\sum_{p \in \cU} |p| \leq W_r/20+|r|$.
\end{corollary}
\begin{proof}
 Let $c_1, \dots, c_m$ be the children of $r$ in $\cU$.  By definition, $W_r = \sum_i W_{c_i}+R^+_r \geq \sum_i W_{c_i}$.
 By \Clm{charge}, for each $c_i$ we have $W_{c_i}/20\geq \sum_{p \in \cU_{c_i}} |p|$.  Combining these to facts yields the claim.
\end{proof}

% \begin{claim} \label{clm:charge} Fix $p \in \cU$. Suppose for any $q,q' \in \cU_p$
% where $q$ is a parent of $q'$ in $\cU_p$, $W_q \geq (4/3)W_{q'}$. Then $\sum_{q \in \cU_p} |q| \leq W_p/20$.
% \end{claim}
% 
% \begin{proof} Since $q$ is an unsupported light path, $|q| < \sqrt{|H_q|}/100 \leq \sqrt{|\redH_q|}/99 \leq \sqrt{|W_q|}/99$.
% We prove that $\sum_{q \in \cU_p} \sqrt{|W_q|}/99 \leq W_p/20$
% by a charge redistribution scheme. Assume that each $q \in \cU_p$
% starts with $\sqrt{W_p}/99$ units of charge. We redistribute this charge over all
% nodes in $\cU_p$, and then calculate the total charge. For $q \in \cU_p$, spread
% its charge to all nodes in $\cU_q$ proportional to $R^+$ values. In other words,
% give $(\sqrt{W_q}/99)\cdot(R^+_{q'}/W_q)$ units of charge to each $q' \in \cU_q$.
% 
% After the redistribution, let us compute the charge deposited at $q$. Every ancestor in $\cU_p$ 
% $q = a_0, a_1, a_2, \ldots, a_k$ contributes to the charge at $q$. The charge is expressed in the following
% equation. We use the assumption that $W_{a_i} \geq (4/3) W_{a_{i-1}}$ and hence $W_{a_i} \geq (4/3)^i W_{a_0} \geq (4/3)^i$, 
% as $a_0$ is an unsupported light path and hence $W_{a_0}\geq 1$. 
% $$ ({R^+_q}/99)\sum_{a_i} 1/\sqrt{W_{a_i}} \leq ({R^+_q}/99)\sum_{a_i} (3/4)^{i/2} \leq R^+_q/20 $$
% The total charge is $\sum_{q \in \cU_p} R^+_p/20 = W_p/20$. 
% \end{proof}

@@

% \begin{proof} We use $r$ for $r(\cU)$. Our aim is to simply apply \Clm{charge} from $r(\cU)$. 
% Suppose $W_q \geq (4/3)W_{q'}$ (for any choice in $\cU$ of $q$ parent of $q'$). By \Clm{charge}, $\sum_{p \in \cU} |p| \leq W_{r}/20$.
% By \Clm{loss}, $|\redH_r| \geq W_r - 2\sum_{p \in \cU} |p|$. By simple algebra, we deduce, $\sum_{p \in \cU} |p| \leq |\redH_r|/10 = |H_r|/10$.
% 
% We now prove that for any $q$ parent of $q'$, $W_q \geq (4/3)W_{q'}$.
% Suppose not. Let $p, p'$ be the counterexample furthest from the root,
% where $p$ is the parent of $p'$. 
% Note that for $q$ and child $q'$ in $\cU_{p'}$, $W_q \geq (4/3)W_{q'}$.
% We will apply \Clm{charge} for $\cU_{p'}$ to deduce that $\sum_{q \in \cU_{p'}} |q| \leq W_{p'}/20$.
% By \Clm{loss}, $|\redH_{p'}| \geq 19W_{p'}/20$. By \Lem{geometric}, $|\redH_p| \geq (3/2)|\redH_{p'}|$.
% Noting that $W_p \geq |\redH_p|$, we deduce that $W_p \geq (4/3)W_{p'}$.
% Hence, $p, p'$ is not a counterexample, and more generally, this is no counterexample.
% That completes the whole proof.
% \end{proof}

\begin{proof} We use $r$ for $r(\cU)$. %Our aim is to simply apply \Clm{charge} from $r(\cU)$. 
Suppose $W_q \geq (4/3)W_{q'}$ (for any choice in $\cU\setminus\{r\}$ of $q$ parent of $q'$), 
then by \Cor{chargeCor}, $\sum_{p \in \cU} |p| \leq W_{r}/20+|r|$.
By \Clm{loss}, $|\redH_r| \geq W_r - 2\sum_{p \in \cU} |p|$, and so combining these inequalities gives,
\[
% \sum_{p \in \cU} |p| \leq \pth{|\redH_r|+2\sum_{p \in \cU} |p|}/20+|r| \Rightarrow 
 \sum_{p \in \cU} |p|\leq \frac{10}{9} \pth{|\redH_r|/20+|r|} \leq 
 \frac{10}{9}\pth{|H_r|/20+\sqrt{|H_r|}/100} \leq |H_r|/10.
\]
